### 2016-04-13 docs/papers/2015eurasip/code/eigen_analysis/html/eigencorrelation.tex

% This LaTeX was auto-generated from MATLAB code.
% To make changes, update the MATLAB code and republish this document.

\documentclass{article}
\usepackage{graphicx}
\usepackage{color}

\sloppy
\definecolor{lightgray}{gray}{0.5}
\setlength{\parindent}{0pt}

\begin{document}

\section*{Input}

\begin{par}
Matrix
\end{par} \vspace{1em}

\subsection*{Contents}

\begin{itemize}
\setlength{\itemsep}{-1ex}
   \item Output
\end{itemize}

\subsection*{Output}

\begin{par}
(R) correlation matrix (E) eigenvalues diagonal matrix (V) eigenvectors (M) largest eigenvalue
\end{par} \vspace{1em}
\begin{verbatim}
function [R,E,V,M] = eigencorrelation(X)
    numLines = size(X,1);
    numColumns = size(X,2);
    Y = zeros(numLines,numColumns);

    for i = 1:numLines
        stdv = std(X(i,:),1);
        if stdv > 0
            Y(i,:) = (X(i,:) - mean(X(i,:))) / stdv;	% standardization, have zero mean and unit variance
        end
    end

    R = 1/numColumns*Y*Y';								% estimation of correlation matrix (R)
    [V,E] = eig(R);										% eigenvectors (V) and eigenvalues (E) of correlation matrix
    M(1) = max(diag(E));                                % largest eigenvalue (M)

    Ed  = diag(E);
    for i = 1:size(Ed,1);
        if M(1) == Ed(i)
            M(2) = i;                                   % get the position of largest eigenvalue
        end
    end
\end{verbatim}

        \color{lightgray} \begin{verbatim}Error using eigencorrelation (line 10)
Not enough input arguments.
\end{verbatim} \color{black}

\end{document}