### 2017-09-19 docs/thesis/thesis.tex

\documentclass[en, oneside, onehalfspacing]{risethesis}

\usepackage{natbib}
\usepackage{babel}
\usepackage{supertabular}
\usepackage{fancybox}
\usepackage{acronym}
\usepackage{array}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{rotating}
\usepackage{tabularx}
\usepackage{colortbl}
\usepackage{multirow}
\usepackage{setspace}
\usepackage{placeins}
\usepackage{longtable}
\usepackage{mathtools}
\usepackage{listings}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{epstopdf}
\usepackage{subfigure}
\lstset{numbers=left,
	stepnumber=1,
	firstnumber=1,
	numberstyle=\tiny,
	extendedchars=true,
	breaklines=true,
	frame=tb,
	basicstyle=\footnotesize,
	stringstyle=\ttfamily,
	showstringspaces=false
}
\renewcommand{\lstlistingname}{Code}
\renewcommand{\lstlistlistingname}{Lista de Listagens}
\definecolor{lightgray}{gray}{0.8}
\usepackage[linkcolor=blue, citecolor=blue, urlcolor=blue, colorlinks, pdfpagelabels, pdftitle={ycc-msc}, pdfauthor={Thiago Pereira de Brito Vieira}]{hyperref}

\address{BRASÍLIA}

\universitypt{Universidade de Brasília}
\universityen{Universidade de Brasília}

\departmentpt{Departamento de Engenharia Elétrica - ENE/FT}
\departmenten{Departamento de Engenharia Elétrica - ENE/FT}

\programpt{Progama de Pós-Graduação em Engenharia Elétrica - PPGEE}
\programen{Progama de Pós-Graduação em Engenharia Elétrica - PPGEE}

\majorfieldpt{Engenharia Elétrica}
\majorfielden{Electrical Engineering}

\title{Discriminative Sensing Based on Signal Processing}
\date{June/2018}

\author{Thiago Pereira de Brito Vieira}
\adviser{Adviser}
\coadviser{coadviser}

\begin{document}

\frontmatter
\frontpage
\presentationpage

\begin{dedicatory} Ficha Catalográfica\end{dedicatory}

\begin{dedicatory} Signatures\end{dedicatory}

\begin{dedicatory} Dedicatory.\end{dedicatory}

\agradecimentos
First and foremost, I would like to thank God for giving me the life, strength, knowledge and opportunity to undertake this research study and to ..

(...)

Thank You!!!

\begin{epigraph}[]{Confucius}
Wherever you go, go with all your heart.
\end{epigraph}

\resumo
\input{resumo}

\abstract
\input{abstract}

\tableofcontents

\makeatletter
\renewcommand{\@thesubfigure}{\thesubfigure:\hskip\subfiglabelskip}
\makeatother
\setcounter{lofdepth}{2}

\listoffigures
\listoftables

\input{acronyms}
\mainmatter

\include{chapters/1_introduction/introduction}
%\include{chapters/2_pca_svm_cfa/pca_svm_cfa}
%\include{chapters/3_eigen_similarity/eigen_similarity}
%\include{chapters/4_mobile/mobile}
\include{chapters/6_conclusion_future_work/conclusion}

\bibliographystyle{natbib}
\addcontentsline{toc}{chapter}{Bibliography}
\bibliography{references}

\clearpage
\addappheadtotoc

\end{document}