### 2017-09-20 docs/thesis/chapters/5_tensor_dl/tensor_dl.tex
compiling version with all chapters

\chapter{Tensor-Based Discriminative Dictionary Learning}
\label{ch:tensor_dl}

Tensor-Based Discriminative Dictionary Learning for Fraud Detection