### 2017-09-20 docs/thesis/thesis.tex

% class
\documentclass[en, oneside, onehalfspacing]{risethesis}

% packages and configurations
\usepackage{enumitem}
\usepackage{lineno, hyperref}
\usepackage[numbers]{natbib}
\usepackage{babel}
\usepackage{supertabular}
\usepackage{fancybox}
\usepackage{acronym}
\usepackage{array}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{rotating}
\usepackage{tabularx}
\usepackage{color, colortbl}
\usepackage{multirow}
\usepackage{hhline}
\usepackage{setspace}
\usepackage{placeins}
\usepackage{longtable}
\usepackage{amsthm,amsmath}
\usepackage{mathtools}
\usepackage{algorithm}% http://ctan.org/pkg/algorithms
\usepackage{algpseudocode}
\usepackage{listings}
\usepackage{epstopdf}
\usepackage{subfigure}
\usepackage{courier}
\usepackage{amsfonts}
\usepackage{morefloats}
\usepackage{lipsum}
\usepackage{cleveref}
\usepackage{hyperref}

\newtheorem{thm}{Scenario}
\hypersetup{breaklinks=true}
\urlstyle{same}
\Urlmuskip=0mu plus 1mu\relax
\lstset{numbers=left, 
	stepnumber=1, 
	firstnumber=1, 
	numberstyle=\tiny, 
	extendedchars=true, 
	breaklines=true,
	frame=tb,
	basicstyle=\footnotesize, 
	stringstyle=\ttfamily,
	showstringspaces=false
}

\definecolor{lightgray}{gray}{0.8}
\definecolor{Gray}{gray}{0.9}
\renewcommand{\lstlistingname}{Code}
\renewcommand{\lstlistlistingname}{Lista de Listagens}

\DeclarePairedDelimiter\abs{\lvert}{\rvert}%
\DeclarePairedDelimiter\norm{\lVert}{\rVert}%
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator*{\argmax}{arg\,max}
\newcommand*\justify{%
	\fontdimen2\font=0.4em% interword space
	\fontdimen3\font=0.2em% interword stretch
	\fontdimen4\font=0.1em% interword shrink
	\fontdimen7\font=0.1em% extra space
	%\hyphenchar\font=`\-% allowing hyphenation
}

% Algorithmic modifications
\makeatletter
\newcommand{\ALOOP}[1]{\ALC@it\algorithmicloop\ #1%
  \begin{ALC@loop}}
\newcommand{\ENDALOOP}{\end{ALC@loop}\ALC@it\algorithmicendloop}
\renewcommand{\algorithmicrequire}{\textbf{Input:}}
\renewcommand{\algorithmicensure}{\textbf{Output:}}
\newcommand{\algorithmicbreak}{\textbf{break}}
\newcommand{\BREAK}{\STATE \algorithmicbreak}
\makeatother

% 
\address{BRASÍLIA}

\universitypt{Universidade de Brasília}
\universityen{Universidade de Brasília}

\departmentpt{Departamento de Engenharia Elétrica - ENE/FT}
\departmenten{Departamento de Engenharia Elétrica - ENE/FT}

\programpt{Progama de Pós-Graduação em Engenharia Elétrica - PPGEE}
\programen{Progama de Pós-Graduação em Engenharia Elétrica - PPGEE}

\majorfieldpt{Engenharia Elétrica}
\majorfielden{Electrical Engineering}

\title{Discriminative Sensing Based on Signal Processing}
\date{June/2018}

\author{Thiago Pereira de Brito Vieira}
\adviser{João Paulo Carvalho Lustosa da Costa}
\coadviser{?}

\begin{document}

\frontmatter
\frontpage
\presentationpage

\begin{dedicatory}Ficha Catalográfica\end{dedicatory}

\begin{dedicatory}Signatures\end{dedicatory}

\begin{dedicatory}Dedicatory.\end{dedicatory}

\agradecimentos
First and foremost, I would like to thank God for giving me the life, strength, knowledge and opportunity to undertake this research study and to ..

(...)

Thank You!!

\begin{epigraph}[]{Confucius}
	Wherever you go, go with all your heart.
\end{epigraph}

\resumo
\input{chapters/resumo}

\abstract
\input{chapters/abstract}

\tableofcontents

\makeatletter
\renewcommand{\@thesubfigure}{\thesubfigure:\hskip\subfiglabelskip}
\makeatother
\setcounter{lofdepth}{2}

\listoffigures
\listoftables

\input{chapters/acronyms}
\mainmatter

\include{chapters/1_introduction/introduction}
\include{chapters/2_pca_svm_cfa/pca_svm_cfa}
\include{chapters/3_eigen_similarity/eigen_similarity}
\include{chapters/4_mobile/mobile}
\include{chapters/5_tensor_dl/tensor_dl}
\include{chapters/6_conclusion_future_work/conclusion}

\bibliographystyle{natbib}
\addcontentsline{toc}{chapter}{Bibliography}
\bibliography{references}

\clearpage
\addappheadtotoc

\end{document}