### 2019-05-27 docs/thesis/chapters/4_m_rpca/m_rpca.tex

\chapter{Moment Distances from Robust Subspace for Network Anomaly Detection}
\label{ch:4_m_rpca}

\begin{quotation}[]{Paulo Freire}
No one knows it all. No one is ignorant of everything. We all know something. We are all ignorant of something.
\end{quotation}

Some widely adopted algorithms for anomaly detection assume a gaussian distributed data, however this assumption may not be observed in some real world problems, such as the case of network traffic analysis, where network traffic features are usually more characterized by skewed and heavy-tailed distributions \cite{lakhina2005mining,benson2010network}.

Findings of Benson \emph{et al.}  \cite{benson2010network} indicate that certain positive skewed and heavy-tailed distributions can model data center switch traffic, and highlights a difference between the data center environment and the wide area network, where the long-tailed Pareto distribution typically shows the best fit \cite{benson2010network}. Leon-Garcia \cite{leon2017probability} also argues that Pareto distribution has been found to capture the behavior of many quantities of interest in the study of Internet behavior. Moreover, Benson \emph{et al.}  \cite{benson2010network} observes that the lognormal distribution is the best fit to model arrival processes in a data center.

This findings show that the skeweness and heavy-tailed distributions may be important for network traffic analysis, and can motivate evaluations to study the impact of skewed data into algorithms that rely on gaussian distributed data, as well as it can highlight opportunities to  exploit the skeweness and heavy-tailed distributions in order to obtain accurate classifiers for network anomaly detection.

Network anomaly detection problems are usually characterized by imbalanced data \cite{Phua2004minority,he2008learning,benson2010network}. However, learning algorithms for imbalanced data has been a challenging research topic, considering that the fundamental issue with the imbalanced learning problem is the ability of imbalanced data to significantly compromise the performance of most standard learning algorithms \cite{he2008learning}. Therefore, learning methods for imbalanced and skewed data have been attracted attention of researchers \cite{Phua2004minority,hubert2009robustskewed}.

We believe that the skewness of anomalous and normal traffic can highlight features for improving anomaly detection in imbalanced data, 
and that the distance between robust estimates of normal traffic and anomalies can highlight discrepancies and be used for network attack detection. Therefore, we propose the md-RPCA, which is an approach based on distances from moments computed from robust subspace learned by Robust Principal Component Analysis, in order to detect anomalies from network traffic. The proposed approach relies on a robust subspace of normal traffic for estimating higher moments. The anomaly detection from new observations evaluate the Mahalanobis distance between the robust higher moments and the higher moments of new observations, in a semi-supervised or unsupervised fashion, without the computational cost of new robust subspace learning for new observations.

We evaluate the accuracy of the md-RPCA for anomaly detection on synthetic data set and on the CTU-13 data set \cite{garcia2014empirical}, which is a large data set of normal, background and botnet traffic that has been adopted to deal with the lack of up-to-date real-world data sets for anomaly detection systems \cite{osanaiye2016distributed}. The Experimental evaluation compares our proposal to widely adopted algorithms for anomaly detection based on clustering and statistical approaches.

% We also evaluate the results of ROBPCA \cite{hubert2005robpca}, which is a method that also relies on robust estimates with adjusted outlyingness based on robust skewness.

The main contribution of this work is a novel semi-supervised and unsupervised method for anomaly detection in skewed and imbalanced data, with results of experimental evaluation on synthetic and real data sets.

This Chapter is organized as follows. In Section \ref{sec:4_relatedworks}, it is conducted a literature review of network anomaly detection, botnet detection, and imbalanced learning. Section \ref{sec:4_datamodel} presents the data model and the evaluated data set. Section \ref{sec:4_proposal} describes the proposed approach for network attack detection. Section \ref{sec:4_experiments} discusses the experimental validation and Section \ref{sec:4_results} presents the results. Finally, Section \ref{sec:4_conclusion} draws the conclusions and the suggestions for future work.

\section{Related Works}
\label{sec:4_relatedworks}

% TODO - re-write this, it is strange
% network anomaly detectiom
Bhuyan \emph{et al.} \cite{bhuyan2014network} provide an overview of facets of network anomaly detection, present attacks normally encountered by network intrusion detection systems, and categorize existing network anomaly detection methods and systems based on the underlying techniques. Ahmed \emph{et al.} \cite{ahmed2016survey} present an analysis of four major categories of anomaly detection techniques which include classification, statistical, information theory and clustering. Moustafa \emph{et al.} \cite{moustafa2019holistic} discuss aspects of anomaly-based Network Intrusion Detection Systems (NIDSs), explains cyber-attacks and new solutions for anomaly detection, and provides a benchmark data sets for training and validating approaches for detection.

% botnet detection
A botnet is a network of bots, that are compromised machines under the influence of malware (bot). The botnet is commandeered by a botmaster and used as resource for attacks, such as distributed denial-of-service (DDoS) attacks, and fraudulent activities such as spam, phishing, identity theft, and information ex-filtration. We refer to \cite{ahmed2016survey} and \cite{moustafa2019holistic} for an overview of network attacks. The botmaster coordinate a botnet through a command and control (C\&C) channel where bots receive commands and synchronize attacks and fraudulent activities. Centralized C\&C structures using the Internet Relay Chat (IRC) protocol have been utilized by botmasters for a long time, but other protocols, such as HTTP, and architectures, such as Peer-To-Peer, have also been adopted \cite{gu2008botminer}.

Acarali \emph{et al.} \cite{acarali2016survey} surveys network-based detection approaches for HTTP-based botnets, surveys traffic-based features used to detect bot traffic and presents an abstraction of the main types of features related to protocols and OSI layers. Wang \emph{et al.} \cite{Wang2018ddosbotnetssurvey} present an analysis based on 50,704 different Internet DDoS attacks originated of 674 botnets from 23 different botnet families with a total of 9,026 victim belonging to 1,074 organizations in 186 countries. Their analysis reveals that geolocation of the attacking sources follows patterns and enables source prediction, and highlights that multiple attacks to the same target also exhibit strong patterns of inter-attack time interval, also presents that there is a trend for different botnets to launch DDoS attacks targeting the same victim, simultaneously or in turn.

The BotHunter was proposed by Gu \emph{et al.} \cite{gu2007bothunter} to detect the infection and coordination of botnets by matching sequence model, through a correlation approach for detecting stages of the infection process.  Gu \emph{et al.} \cite{gu2008botminer} presented the BotMiner, which aims to detect groups of compromised machines that are part of a botnet. BotMiner monitors communications that may suggest C\&C or malicious activities, and finds a coordinated group pattern by means of clusters of similar communication activities, clusters of similar malicious activities, and performs cross cluster correlation to identify the hosts that share both similar communication patterns and similar malicious activity patterns.

Khattak \emph{et al.} \cite{khattak2015botflex} proposed BotFlex, which is a network-based tool for botnet detection, composed by a Complex Event Processing (CEP) engine and on a correlation framework that continuously receives events and correlates them according to rules. BotFlex's results are compared to BotHunter \cite{gu2007bothunter}, but the evaluation relies in a own and not public data set.

% justification for the data set and present who else use it
Several approaches for network attack detection uses the KDD 99 \cite{ahmed2016survey,osanaiye2016distributed,bhuyan2014network} data sets for accuracy and performance evaluation, due to their availability and labeled attacks. Even though the KDD 99 data set are criticized by the generation procedure and the risk of over-estimations of anomaly detection due to data redundancy, it still represents one of the few publicly available labeled data sets currently in use today by researchers \cite{osanaiye2016distributed,bhuyan2014network}. NSL-KDD \cite{tavallaee2009detailed} data set is the refined version of the KDD 99 data set that redundant data records are removed, in order to avoid biased classifications. However, NSL-KDD data set maintain the limitations of the KDD 99 regarding volume and reproduction of network traffic similar to the real networks.

The objective of Garcia \emph{et al.} \cite{garcia2014empirical} is to compare three botnet detection methods using a simple and reproducible methodology, a good data set and a new error metric. This paper evaluates some data sets for network anomaly detection, surveys some approaches for botnet detection, proposes two methods (BClus and CAMNEP) for botnet detection and compare results to BotHunter \cite{gu2007bothunter}.

Considering the lack of available labeled data sets, Garcia \emph{et al.} \cite{garcia2014empirical} proposes the CTU-13 data set, which is composed by attack, normal and background labeled data, in an imbalanced distribution like in a real network. The authors recommend scenarios for training and testing, in the way that none of the botnet families used in the training and cross-validation data set should be used in the testing data set, aiming to ensure that the evaluated methods can generalize and detect new behaviors. Adopting the training and testing approach proposed by Garcia \emph{et al.}, some botnet malwares wouldn't be tested, since in the author's proposal some botnets are present only for training.

Wang and Paschalidis \cite{wang2017botnet} proposed a botnet detection approach based on anomaly and community detection, aiming for detecting botnets and identifying bots before the botnet becomes active. The first stage detects anomalies by leveraging large deviations of an empirical distribution. The second stage detects the bots using ideas from social network community detection in a graph that captures correlations of interactions among nodes over time. This work is compared with the BotHunter \cite{gu2007bothunter} on the CTU-13 botnet data set \cite{garcia2014empirical}.

% \cite{da2018online} proposes an approach using the Very Fast Decision Tree, a classification algorithm used on stream mining that can learn incrementally when needed, to identify botnets by observing network flows. When evaluating the approach on multiple scenarios with different botnets, we were able to achieve high performance metrics on the majority of scenarios, while using a significantly low number of labelled instances. It discard scenarios 5 and 6. They does not compute F1, but it is easy to compute it from precision and recall provided.

% Anomaly detection by subspace larning
Traditional PCA-based anomaly detection models are not suitable for anomaly interpretation, as they judge whether a data instance is an anomaly or not based on the length of its projection on the abnormal subspace spanned by the less significant principal components, and there is no direct mapping between PCA’s dimensionality-reduced subspace and the original feature space \cite{ringberg2007sensitivity}. However, to overcome the above mentioned limitations, some approaches based on PCA have been proposed for network anomaly detection. Callegari \emph{et al.} \cite{callegari2011novel} proposed a PCA-based method for identifying the network traffic flows responsible for an anomaly detected at the aggregate level, by means of a separation of normal and anomalies according to principal components (normal) and remaining (anomalies). Lee \emph{et al.} \cite{Lee2013} presented OverSampling PCA (osPCA), which allows one to determine the anomaly of the target instance according to the variation of the resulting dominant eigenvector obtained by similarity analysis and over sampling. Vieira \emph{et al.} \cite{vieira2017model} proposed a framework that applies Model Order Selection (MOS) for detection of time frames under attack and uses similarity analysis to extract details and detect the time and ports under attack.

% Anomaly detection by robust subspace learning
The problem of PCA or subspace learning for outlier corrupted data is called Robust Principal Component Analysis (RPCA) or robust subspace learning \cite{candes2011robust, vaswani2018robust}. RPCA aims to be resilient to outliers by means of a robust subspace learning \cite{vaswani2018robust} for outlier corrupted data, decomposing a given data matrix $\textbf{X}$ into the sum of a low rank matrix $\textbf{L}$, whose column subspace gives the principal components, and a sparse matrix $\textbf{S}$, which refers to outliers’ matrix. This definition is also referred to as the sparse + low rank (S + LR) formulation \cite{vaswani2018robust}. We refer to \cite{lerman2018overview} and \cite{vaswani2018robust} for more details regarding robust subspace learning.

RPCA is been mainly applied to computer vision, in problems of robust subspace tracking and robust subspace recovery. However, RPCA has also been adopted for general outlier detection \cite{hubert2005robpca,hubert2009robustskewed,cherapanamjeri2017thresholding,zhou2017anomaly,NetflixSurus} and for anomaly detection on network traffic \cite{pascoal2012robust}. ROBPCA \cite{hubert2005robpca} intends to identify outliers using PCA from robust estimates of mean and covariance matrix, to reduce the data dimensions and plotting the orthogonal distances versus the robust score distances, to flag an outlier map. However, ROBPCA flags many points as outlying when the original data is skewed. Therefore, Hubert \emph{et al.} \cite{hubert2009robustskewed} proposed ROBPCA-AO, which improves ROBPCA for problems with skewed data, by means of an adjusted outlyingness based on robust skewness.

Robust subspace learning has receiving a growing attention of researchers aiming the development of network anomaly detection systems \cite{rousseeuw1984mcd, rousseeuw1999fastmcd, hubert2005robpca,hubert2009robustskewed, pascoal2012robust, zhou2017anomaly}, considering outlier-robust methods and sparse-corruption methods \cite{lerman2018overview}. Pascoal \emph{et al.} \cite{pascoal2012robust} proposed an approach based on a robust mutual information estimator for feature selection and based on RPCA for outlier detection in internet traffic. The anomaly detection proposed by Pascoal \emph{et al.} is an unsupervised approach that estimate the first $k$ robust principal components, calculate the score and the orthogonal distances, calculate the thresholds and classify new observations accordingly. Zhou and Paffenroth \cite{zhou2017anomaly} proposed the Robust Deep Autoencoders (RDA), which the central idea is that a RDA inherits the non-linear representation capabilities of autoencoders combined with the anomaly detection capabilities of RPCA. Considering that outliers and noise may reduce the quality of representations discovered by deep autoencoders, the proposed model isolates noise and outliers in the input by means of a RPCA approach, and the autoencoder is trained after this isolation. RDA was evaluated by the authors for the MNIST data set.

Benson \emph{et al.} \cite{benson2010network} conducted an empirical study of the network traffic in 10 data centers belonging to three different types of organizations, including university, enterprise, and cloud data centers. Findings of Benson \emph{et al.} indicate that certain positive skewed and heavy-tailed distributions can model data center switch traffic, and highlights a difference between the data center environment and the wide area network, where the long-tailed Pareto distribution typically shows the best fit \cite{benson2010network}.

Mahalanobis Distance (MD) is a generalized distance which is useful for determining the similarity between an unknown sample and a collection of known samples, by considering the covariances between the variables and their mean values. The MD is a measure of the distance between a vector $\boldsymbol{x}$ and a distribution $\boldsymbol{X}$, introduced by P. C. Mahalanobis in 1936 \cite{mahalanobis1936md}. It is a multi-dimensional generalization for measuring how many standard deviations away $\boldsymbol{x}$ is from the mean $\boldsymbol{\bar{x}}$ and covariance $\boldsymbol{\bar{\Sigma}}$ of $\boldsymbol{X}$. MD has been used for distance based anomaly detection with robust estimates in many areas, assuming that the Mahalanobis Distance between robust estimates and new observations can reveal anomalies.

We propose an approach based on higher moments distances from robust subspace for botnet detection. The proposed approach relies on a robust subspace of normal traffic for estimating higher moments. Thus, we propose that the anomaly detection for new observations should evaluate the Mahalanobis distance between the robust higher moments and the higher moments of new observations, in a semi-supervised fashion and without the computational cost of new robust subspace learning for new observations, or in a unsupervised approach.

\section{Data Model}
\label{sec:4_datamodel}

In this paper, scalars are denoted by italic letters (\emph{a, b, A, B, $\alpha$, $\beta$}), vectors by lowercase bold letters (\textbf{a, b}), matrices by uppercase bold letters (\textbf{A, B}), and $a_i,_j$ denotes the (\emph{i, j}) elements of the matrix \textbf{A}. The superscripts \textsuperscript{T} and \textsuperscript{-1} are used for matrix transposition and matrix inversion, respectively. The Frobenius norm is denoted as $\left\| \mathord{\cdot} \right\|_F$, while $\left\| \mathord{\cdot} \right\|_*$ denotes the nuclear norm of a matrix and $\left\| \mathord{\cdot} \right\|_1$ means the sum of the absolute values of matrix entries. We also define the operator $\langle \mathord{\cdot} \rangle$, that is the standard trace inner product, and $[ \mathord{\cdot} ]^n$, which denotes the $n$ largest values of a vector.

This section also presents a description of the synthetic data model in Subsection \ref{sec:4_synthetic} and in Subsection \ref{sec:4_CTU-13} we describe the data model for the CTU-13 data set.

\subsection{The Synthetic data set}
\label{sec:4_synthetic}

In order to evaluate the null hypothesis $H_{0rob.sub}$, which assumes that the robust subspace learning does not improves the anomaly detection from imbalanced and skewed, we create two synthetic data sets characterized by skewed and heavy tailed distributions. We selected Pareto and Lognormal distributions to evaluate the anomaly detection capacity of our proposal, considering that these distributions have been adopted to represent network traffic in the internet and data centers \cite{benson2010network,leon2017probability}.

We adopt a contamination rate of 33\% to reproduce the imbalanced data of anomaly detection problems, and select the gaussian distribution for the addictive anomalies into Pareto and Lognormal distributions. Therefore, we generate $\boldsymbol{Y}_p$ and $\boldsymbol{Y}_l$ with 2,400 samples and two features, where 793 samples are gaussian anomalies, also referred as noise or contamination, while 1,607 samples are legitimate signals. The data set $\boldsymbol{Y}_p$ is composed by random Pareto distribution with random gaussian noise, while $\boldsymbol{Y}_l$ is composed by random Lognormal distribution with random gaussian noise.

The Figure \ref{fig:4.03} shows the histogram and the PDF of the Pareto distribution for two features of $\boldsymbol{Y}_p$.

\begin{figure}[!htb]
	\centering
	\subfigure[Feature 1]{%
		\includegraphics[width=0.48\textwidth]{figures/pareto1.png}
		\label{fig:4.01}
	}
	\subfigure[Feature 2]{%
		\includegraphics[width=0.48\textwidth]{figures/pareto2.png}
		\label{fig:4.02}
	}
	\caption[Pareto Distribution]{Pareto Distribution}
	\label{fig:4.03}
\end{figure}

The Figure \ref{fig:4.06} shows the histogram and the PDF of the Lognormal distribution for two features of $\boldsymbol{Y}_l$.

\begin{figure}[!htb]
	\centering
	\subfigure[Feature 1]{
		\includegraphics[width=0.48\textwidth]{figures/lognormal3.png}
		\label{fig:4.04}
	}
	\subfigure[Feature 2]{
		\includegraphics[width=0.48\textwidth]{figures/lognormal4.png}
		\label{fig:4.05}
	}
	\caption[Lognormal Distribution]{Lognormal Distribution}
	\label{fig:4.06}
\end{figure}

The Pareto distribution contaminated by gaussian noise is denoted as $\boldsymbol{Y}_p^c$, while $\boldsymbol{Y}_l^c$ denotes the Lognormal distribution contaminated by gaussian noise. Taking into account that semi-supervised algorithms requires training from normal data, we also generate $\boldsymbol{Y}_p^t$, while $\boldsymbol{Y}_l^t$, which are data sets of Pareto and Lognormal distributions for training, without contamination and composed by 1,607 samples and two features.

\subsection{The CTU-13 data set}
\label{sec:4_CTU-13}

The CTU-13 is a data set of botnet traffic that was captured in the Czech Technical University \cite{garcia2014empirical}, by means of a testbed and execution of malwares in a real network. The CTU-13 data set contains 13 scenarios with network flows of botnet malwares, that are: neris, rbot, virut, menti, sogou, nsys.ay and murlo. The botnet traffic is also classified as attack or command and control (C\&C), while the legitimate flows can also be classified as normal or background. It is important to note that the CTU-13 data set can also be formulated as a superposition of signal, artefact and noise, which refer to background, normal a botnet traffic, respectively.

The types of C\&C and attack flows present in CTU-13 data set are:

\begin{itemize}
	\item \textbf{Attacks:} Click Fraud (CF), Port Scan (PS), Compiled and Controlled by Authors (US), SPAM and DDOS;
	\item \textbf{C\&C:} IRC, P2P and HTTP.
\end{itemize}

We refer to Garcia \cite{garcia2014identifying} and Garcia \emph{et al.} \cite{garcia2014empirical} for a detailed description of the performed attacks and C\&C flows, also including the topology of the adopted testbed, rules for classifying normal flows, and an analysis of behaviors or patterns of the malware's traffic.

For all the scenarios, the authors of the CTU-13 data set convert the captured pcap files to NetFlows and release the processed flows. The data set contains ground-truth labels for flows as follows: flows from or to the infected machines are labeled as “botnet”; flows from or to well known and controlled machines are labeled as “normal”; all other flows are labeled as “background.”

Table \ref{tab:4.01} presents an overview organized by scenario and shows the malwares used for botnet attacks, the types of attacks and C\&C types, the total number of flows, the number of malicious flows, which includes flows of C\&C and attacks, and finally shows the number of normal flows.

\begin{table}[h!]
	\scriptsize
	\caption{CTU-13 data set Description}
	\label{tab:4.01}
	\begin{tabular}{| l | l | l | r | r | r | r | r | r | r | r | }
		\hline \rowcolor{Gray} \begin{tabular}[x]{@{}l@{}}\textbf{ID}\end{tabular}	&\begin{tabular}[x]{@{}l@{}}\textbf{Bot}\end{tabular}	 &\begin{tabular}[x]{@{}l@{}}\textbf{Type}\end{tabular}	&\begin{tabular}[x]{@{}l@{}}\textbf{Total}\end{tabular} &\begin{tabular}[x]{@{}l@{}}\textbf{Malicious}\end{tabular} &\begin{tabular}[x]{@{}l@{}}\textbf{C\&C}\end{tabular} &\begin{tabular}[x]{@{}l@{}}\textbf{Attack}\end{tabular} &\begin{tabular}[x]{@{}l@{}}\textbf{Normal}\end{tabular}\\ \hline
			10 & neris &\makecell[l]{IRC, Spam,\\CF} & 2,824,636 & 40,961 (1.45\%) & 341 (0.01\%) & 40,620 (1.44\%) &30,387 (1.07\%)\\ \hline
			11 & neris &\makecell[l]{IRC, Spam,\\CF} & 1,808,122 & 20,941 (1.16\%) & 673 (0.04\%) & 20,268 (1.12\%) &9,120 (0.5\%)\\ \hline
			12 & rbot &\makecell[l]{IRC, PS,\\US} & 4,710,638 & 26,822 (0.57\%) & 63 (0.00\%) & 26,759 (0.57\%) &116,887 (2.48\%)\\ \hline
			15 & rbot &\makecell[l]{IRC, DDoS,\\US} & 1,121,076 & 2,580 (0.23\%) & 52 (0.00\%) & 2,528 (0.23\%) &25,268 (2.25\%)\\ \hline
			15-2 & virut &\makecell[l]{Spam, PS,\\HTTP} & 129,832 & 901 (0.69\%) & 24 (0.02\%) & 877 (0.68\%) &4,679 (3.6\%)\\ \hline
			16 & menti &PS & 558,919 & 4,630 (0.83\%) & 199 (0.04\%) & 4,431 (0.79\%) &7,494 (1.34\%)\\ \hline
			16-2 & sogou &HTTP & 114,077 & 63 (0.06\%) & 26 (0.02\%) & 37 (0.03\%) &1,677 (1.47\%)\\ \hline
			16-3 & murlo &PS & 2,954,230 & 6,127 (0.21\%) & 1,074 (0.04\%) & 5,053 (0.17\%) &72,822 (2.46\%)\\ \hline
			17 & neris &\makecell[l]{IRC, Spam,\\CF, PS} & 2,087,508 & 184,987 (8.86\%) & 2,973 (0.14\%) & 182,014 (8.72\%) &43,340 (1.57\%)\\ \hline
			18 & rbot &\makecell[l]{IRC, DDoS,\\US} & 1,309,791 & 106,352 (8.12\%) & 33 (0.00\%) & 106,319 (8.12\%) &15,847 (1.2\%)\\ \hline
			18-2 & rbot &\makecell[l]{IRC, DDoS,\\US} & 107,251 & 8,164 (7.61\%) & 2 (0.00\%) & 8,162 (7.61\%) &2,718 (2.53\%)\\ \hline
			19 & nsys.ay &P2P & 325,471 & 2,168 (0.67\%) & 25 (0.01\%) & 2,143 (0.66\%) &7,628 (2.35\%)\\ \hline
			15-3 & virut &\makecell[l]{Spam, PS,\\HTTP} & 1,925,149 & 40,003 (2.08\%) & 536 (0.03\%) & 39,467 (2.05\%) &31,939 (1.65)\\ \hline
	\end{tabular}
\end{table}

The CTU-13 data set originally contains the following features:

\begin{itemize}
	\item Start Time;
	\item End Time;
	\item Duration;
	\item Source IP Address;
	\item Source Port;
	\item Direction;
	\item Destination IP Address;
	\item Destination Port;
	\item State;
	\item Type of service from source to destination;
	\item Type of service from destination to source;
	\item Total of Packets;
	\item Total of Bytes;
\end{itemize}

Our analysis of the available features leads to discard some features, considering that highly correlated features can bias or not improve the model, and that source or destination IP addresses can insert some false bias into learning models, since they can be changed by IP spoofing or other techniques.

TODO: talk about the feature distribution, that arent similar to pareto or lognormal, but it is not normal too

An Exploratory Data Analysis (EDA) reveals that distributions of some features are skewed and present an overlapping between normal and anomalous flows, as can be seen in  \ref{fig:4.01}, that presents the distribution of transaction state of scenario 16, and exemplifies the data skewness, the overlapping and the long tail distribution of normal data. 
\begin{figure}[h!]
     \centering
     \includegraphics[width=10cm]{figures/raw_distplot_capture20110816_State.png}
     \caption{Distribution of flow duration for scenario 12.}
     \label{fig:4.01}
\end{figure}

Due to the number of available features and the class overlapping, we performed an correlation analysis and cross validation in order to identify the best set of features for network attack detection. Therefore, we adopt the following features: State, Type of service from source to destination, Type of service from destination to source, destination port, source port, and the total number of packets.

The full data set and scenarios can be denoted as $\boldsymbol{X} = \{\boldsymbol{X}_{10}, \boldsymbol{X}_{11}, \ldots , \boldsymbol{X}_{18-2}, \boldsymbol{X}_{19}\}$, in accordance to IDs presented in Table \ref{tab:4.01}. Each contaminated scenario $\boldsymbol{X}_i$ is split into $\boldsymbol{X}_i^n$ containing 50\% of the the normal data, and into $\boldsymbol{X}_i^t$ that is composed by all anomalous flows and ne needed number of normal flows to have a testing data with 33\% of contamination.

\section{Moment Distances from Robust Subspace for Botnet Detection}
\label{sec:4_proposal}

This section describes the proposed approach for network attack detection by means of a distance analysis between moments computed from a robust subspace and new observations of network traffic.

Robust subspace learning can be defined as the decomposition of a given data matrix $\textbf{X}$ into the sum of a low rank matrix $\textbf{L}$, whose column subspace gives the principal components, and a sparse matrix $\textbf{S}$, with outliers or noise. We propose to learn a robust subspace $\textbf{L}$ from the normal traffic $\textbf{X}^r$ for computing $\textbf{L}^r$, $\textbf{S}^r$ and the robust moments, i.e. the mean $\boldsymbol{\bar{x}}$, skewness $\boldsymbol{\bar{s}}$ and kurtosis $\boldsymbol{\bar{k}}$, in order to evaluate the distance $\boldsymbol{d}$ of new observations $\boldsymbol{X}^t$ to these moments and classify as anomalous the observations with the largest distances.

Robust PCA is a well known method to recover a low-rank matrix $\textbf{L}$ and sparse matrix $\textbf{S}$ from corrupted measurements modeled as $\textbf{X} = \textbf{L} + \textbf{S}$. This decomposition in low-rank and sparse matrices can be achieved by techniques such as Principal Component Pursuit method (PCP), and by optimization methods, such as the Augmented Lagrange Multiplier Method (ALM), Alternating Direction Method (ADM), Fast Alternating Minimization (FAM) or Iteratively Reweighted Least Squares (IRLS) \cite{candes2011robust,vaswani2018robust,lerman2018overview}.

According to Wright et al. \cite{wright2009robust}, under rather broad conditions, as long as the error matrix S is sufficiently sparse, it is possible to recover a low-rank matrix by solving the following convex optimization problem:
\begin{equation}\label{eq:4.01}
	(\boldsymbol{\hat{L}}, \boldsymbol{\hat{S}})\leftarrow \min_{L,S}\left \| \boldsymbol{L} \right \|_{*} + \lambda \left \| \boldsymbol{S} \right \|_{1}
\end{equation}
\begin{center} subject to: $\boldsymbol{X} = \boldsymbol{L} + \boldsymbol{S}$ \end{center}
\begin{equation}\label{eq:4.02}
    \left \| \boldsymbol{L} \right \|_{*} = \sum_{i} \sigma_{i}(\boldsymbol{L})
\end{equation}
\begin{equation}\label{eq:4.03}
    \left \| \boldsymbol{S} \right \|_{1} = \sum_{ij} \left | \boldsymbol{S}_{ij} \right |
\end{equation}
where $\lambda$ is a positive weighting parameter, which determines the sparsity of $\boldsymbol{S}$, $\sigma$ denotes the singular values of a matrix.

Before ALM, some methods were proposed to solve that convex optimization problem, such as Iterative Thresholding (IT) and Accelerated Proximal Gradient (APG). However, according to \cite{lin2010augmented}, both approaches have scalability problems and require a large number of iterations to converge. The Augmented Lagrange Multiplier is proven to have a \emph{Q-linear} convergence speed and experimental results show that ALM is five times faster than APG, which in theory is sub-linear \cite{lin2010augmented}. Furthermore, ALM reaches more accurate results with less iterations.

The RPCA with ALM can be formulated:
\begin{equation}\label{eq:4.04}
	l(\boldsymbol{L}, \boldsymbol{S}, \boldsymbol{Y}) = \left\|\boldsymbol{L}\right\|_* + \lambda\left\|\boldsymbol{S}\right\|_1 + \langle \boldsymbol{Y, X - L - S}  \rangle + \frac{\mu}{2}\left\|\boldsymbol{X - L - S}\right\|_F^2,
\end{equation}
where $\boldsymbol{Y}$ is the multiplier of the linear constraint, $\mu$ is the penalty parameter for the violation of the linear constraint \cite{yuan2009sparse}. Thus, an iterative scheme can be presented as:
\begin{equation}\label{eq:4.05}
	\left\{
		\begin{matrix} 
			(\boldsymbol{L}_{k+1}, \boldsymbol{S}_{k+1}) \in \operatorname*{argmin}_{\boldsymbol{L,S} \in \mathbb{R}^{m \times n}} \{l(\boldsymbol{L}, \boldsymbol{S}, \boldsymbol{Y}_{k})\}, \\ 
			\boldsymbol{Y}_{k+1} = \boldsymbol{Y}_{k} + \mu(\boldsymbol{X} - \boldsymbol{L}_{k} - \boldsymbol{S}_{k}),
		\end{matrix}
	\right.
\end{equation}

We adopt RPCA with ADM, which, generally speaking, is a practical improvement of the classical Augmented Lagrangian method for solving convex programming problem with linear constraints, by fully taking advantage of its high-level separable structure \cite{yuan2009sparse}. ADM minimizes $\boldsymbol{L}$ and $\boldsymbol{S}$ variables serially by solving the following problems to generate the new iterate:

\begin{equation}\label{eq:4.06}
	\left\{\begin{matrix}
	\boldsymbol{L}_{k+1} \in \operatorname*{argmin}_{\boldsymbol{L} \in \mathbb{R}^{m \times n}}\{l(\boldsymbol{L}, \boldsymbol{S}_{k}, \boldsymbol{Y}_{k})\}\\ 
	\boldsymbol{S}_{k+1} \in \operatorname*{argmin}_{\boldsymbol{S} \in \mathbb{R}^{m \times n}}\{l(\boldsymbol{L}_{k+1}, \boldsymbol{S}, \boldsymbol{Y}_{k})\}\\ 
	\boldsymbol{Y}_{k+1} = \boldsymbol{Y}_{k} + \mu(\boldsymbol{X} - \boldsymbol{L}_{k} - \boldsymbol{S}_{k})
	\end{matrix}\right.
\end{equation}

Moments are a set of statistical parameters to measure a distribution. The arithmetic mean is the first general moment, the second is the variance, while skewness (asymmetry) is the third moment and kurtosis (excess) is the fourth moment.

Let the mean $\bar{\boldsymbol{x}} \in \mathbb{R}^{1 \times p}$ be

\begin{equation}\label{eq:4.07}
	\bar{\boldsymbol{x}} = \displaystyle\frac{1}{h}\displaystyle\sum_{i\in \boldsymbol{h}} \boldsymbol{x}_i, 
\end{equation}

and the covariance matrix $\boldsymbol{\bar{\Sigma}} \in \mathbb{R}^{p \times p}$ be

\begin{equation}\label{eq:4.08}
	\boldsymbol{\bar{\Sigma}} = \displaystyle\frac{1}{h}\displaystyle\sum_{i\in \boldsymbol{h}} (\boldsymbol{x}_i - \bar{\boldsymbol{x}})(\boldsymbol{x}_i - \bar{\boldsymbol{x}})^\prime,
\end{equation}

The general formula of the $u$-th moment can be expressed as
\begin{equation}\label{eq:4.09}
	m_u = \displaystyle\frac{1}{n}\displaystyle\sum_{i = 1}^{n}( x_i - \bar{x})^u.
\end{equation}

Therefore, the skewness is calculated by
\begin{equation}\label{eq:4.10}
	\boldsymbol{\bar{s}} = \frac{m_3}{m_2^{\frac{3}{2}}},
\end{equation}

and the kurtosis as
\begin{equation}\label{eq:4.11}
	\boldsymbol{\bar{k}} = \frac{m_4}{m_2^2}.
\end{equation}

We propose to compute the mean $\boldsymbol{\bar{x}}$, the skewness $\boldsymbol{\bar{s}}$, the kurtosis $\boldsymbol{\bar{k}}$ and the covariance matrix $\boldsymbol{\bar{\Sigma}}$ from the robust subspace $\boldsymbol{L}^r$, after to minimize the Equation \ref{eq:4.06} for $\textbf{X}^r$. We also propose to compute the Mahalanobis Distance (MD) for detecting anomalies between new observations and the moments calculated from a robust subspace computed by RPCA.

The classical Mahalanobis Distance is defined as		
\begin{equation}\label{eq:4.12}
	\boldsymbol{d}(\boldsymbol{x},\bar{\boldsymbol{x}}, \boldsymbol{\bar{\Sigma}}) = \sqrt{(\boldsymbol{x} - \bar{\boldsymbol{x}}) \boldsymbol{\bar{\Sigma}}^{-1}(\boldsymbol{x} - \bar{\boldsymbol{x}})^\prime},
\end{equation}
where $\boldsymbol{x}$ is a vector of a new observation, $\bar{\boldsymbol{x}}$ is the mean vector of known observations, also referred as location, and $\boldsymbol{\bar{\Sigma}}$ is the covariance matrix of known observations, also referred as scatter. The classical MD usually relies on robust mean and robust covariance matrix for outlier detection, which are commonly computed by Fast-MCD \cite{rousseeuw1984mcd, rousseeuw1999fastmcd}. Here we propose to calculate the MD based on mean and covariance matrix computed from a robust subspace learned by RPCA. We also propose to extend the Equation \ref{eq:4.12} to implement a Skewness-based Mahalanobis Distance, as follows:

\begin{equation}\label{eq:4.13}
	\boldsymbol{d}(\boldsymbol{x}, \bar{\boldsymbol{s}}, \boldsymbol{\bar{\Sigma}}) = \sqrt{(\boldsymbol{x} - \bar{\boldsymbol{s}}) \boldsymbol{\bar{\Sigma}}^{-1}(\boldsymbol{x} - \bar{\boldsymbol{s}})^\prime}.
\end{equation}

We finally propose to extend the Equation \ref{eq:4.12} to implement a Kurtosis-based Mahalanobis Distance, as follows:

\begin{equation}\label{eq:4.14}
	\boldsymbol{d}(\boldsymbol{x}, \bar{\boldsymbol{k}}, \boldsymbol{\bar{\Sigma}}) = \sqrt{(\boldsymbol{x} - \bar{\boldsymbol{k}}) \boldsymbol{\bar{\Sigma}}^{-1}(\boldsymbol{x} - \bar{\boldsymbol{k}})^\prime}.
\end{equation}

The distances $\boldsymbol{d}(\boldsymbol{x},\bar{\boldsymbol{x}}, \boldsymbol{\bar{\Sigma}})$, $\boldsymbol{d}(\boldsymbol{x}, \bar{\boldsymbol{s}}, \boldsymbol{\bar{\Sigma}})$ and $\boldsymbol{d}(\boldsymbol{x}, \bar{\boldsymbol{k}}, \boldsymbol{\bar{\Sigma}})$ shall be computed and evaluated separately for botnet detection. Therefore, the robust subspace learning from normal data combined to MD can be called md-RPCA when using mean-based MD, sd-RPCA when using skewness-based MD, and kd-RPCA when using kurtosis-based MD.

The contamination rate parameter is widely adopted for well established outlier detection algorithms \cite{zhao2019pyod}, and refers to the percentage of observations that are classified as anomalous. The contamination rate can be well known for some areas, or can be computed by cross-validation or can be assumed according to previous observations. In our proposal, the contamination define the rate of highest distances $\boldsymbol{d}$ that shall be classified as anomalous. Therefore, the resulting $\boldsymbol{\bar{t}}$ refers to observations classified as anomalous according to the contamiation $c$.

The Algorithm \ref{alg:4.01} presents how the m-RPCA can be algorithmically described.
\begin{algorithm}
	\label{alg:4.01}
	\SetAlgoLined
	\KwResult{$\boldsymbol{\bar{t}}_{\bar{\boldsymbol{x}}}$, $\boldsymbol{\bar{t}}_{\bar{\boldsymbol{s}}}$, $\boldsymbol{\bar{t}}_{\bar{\boldsymbol{k}}}$}
	Given $\boldsymbol{X}$ split into $\boldsymbol{X}^n$ and $\boldsymbol{X}^t$\;
	\While{not $\min_{L,S}\left \| \boldsymbol{L} \right \|_{*} + \lambda \left \| \boldsymbol{S} \right \|_{1}$ of $\boldsymbol{X}^n$}{
		$\boldsymbol{L}_{k+1} \in \operatorname*{argmin}_{\boldsymbol{L} \in \mathbb{R}^{m \times n}}\{l(\boldsymbol{L}, \boldsymbol{S}_{k}, \boldsymbol{Y}_{k})\}$\;
		$\boldsymbol{S}_{k+1} \in \operatorname*{argmin}_{\boldsymbol{S} \in \mathbb{R}^{m \times n}}\{l(\boldsymbol{L}_{k+1}, \boldsymbol{S}, \boldsymbol{Y}_{k})\}$\;
		$\boldsymbol{Y}_{k+1} = \boldsymbol{Y}_{k} + \mu(\boldsymbol{X} - \boldsymbol{L}_{k} - \boldsymbol{S}_{k})$\;
	}	
	$\bar{\boldsymbol{x}} = \displaystyle\frac{1}{h}\displaystyle\sum_{i\in \boldsymbol{h}} \boldsymbol{x}_i$, for $\boldsymbol{L}^n$\;
	$\boldsymbol{\bar{\Sigma}} = \displaystyle\frac{1}{h}\displaystyle\sum_{i\in \boldsymbol{h}} (\boldsymbol{x}_i - \bar{\boldsymbol{x}})(\boldsymbol{x}_i - \bar{\boldsymbol{x}})^\prime$, for $\boldsymbol{L}^n$\;
	$\boldsymbol{\bar{s}} = \frac{m_3}{m_2^{\frac{3}{2}}}$, for $\boldsymbol{L}^n$\;
	$\boldsymbol{\bar{k}} = \frac{m_4}{m_2^2}$, for $\boldsymbol{L}^n$\;
	$\boldsymbol{d}(\boldsymbol{x},\bar{\boldsymbol{x}}, \boldsymbol{\bar{\Sigma}}) = \sqrt{(\boldsymbol{x} - \bar{\boldsymbol{x}}) \boldsymbol{\bar{\Sigma}}^{-1}(\boldsymbol{x} - \bar{\boldsymbol{x}})^\prime}$\;
	$\boldsymbol{d}(\boldsymbol{x}, \bar{\boldsymbol{s}}, \boldsymbol{\bar{\Sigma}}) = \sqrt{(\boldsymbol{x} - \bar{\boldsymbol{s}}) \boldsymbol{\bar{\Sigma}}^{-1}(\boldsymbol{x} - \bar{\boldsymbol{s}})^\prime}$\;
	$\boldsymbol{d}(\boldsymbol{x}, \bar{\boldsymbol{k}}, \boldsymbol{\bar{\Sigma}}) = \sqrt{(\boldsymbol{x} - \bar{\boldsymbol{k}}) \boldsymbol{\bar{\Sigma}}^{-1}(\boldsymbol{x} - \bar{\boldsymbol{k}})^\prime}$\;
	$\boldsymbol{\bar{t}}_{\bar{\boldsymbol{x}}} = [\boldsymbol{d}(\boldsymbol{x}, \bar{\boldsymbol{x}}, \boldsymbol{\bar{\Sigma}})]^c$\;
	$\boldsymbol{\bar{t}}_{\bar{\boldsymbol{s}}} = [\boldsymbol{d}(\boldsymbol{x}, \bar{\boldsymbol{s}}, \boldsymbol{\bar{\Sigma}})]^c$\;
	$\boldsymbol{\bar{t}}_{\bar{\boldsymbol{k}}} = [\boldsymbol{d}(\boldsymbol{x}, \bar{\boldsymbol{k}}, \boldsymbol{\bar{\Sigma}})]^c$\;
	\caption{Moment Distances from Robust Subspace}
\end{algorithm}

The steps between 1 and 10 of the Algorithm \ref{alg:4.01} are the training from normal data $\boldsymbol{X}^n$, while the steps between 11 and 16 aim the anomaly detection on new contaminated observations. Note that the steps 11 and 14 refers to the md-RPCA, while the steps 12 and 15 refers to the sd-RPCA, and the steps 13 and 16 refers to the kd-RPCA. It is also important to observe that the anomaly detection from new observations does not require new robust subspace learning, it is only necessary to compute a moment-based Mahalanobis distance to classify the $c$ largest distances as anomalous.

However, m-RPCA can also be adopted as unsupervised algorithm through changing the lines 1 and 2 of the Algorithm \ref{alg:4.01} to substitute $\boldsymbol{X}^n$ by $\boldsymbol{X}^t$. Hence, the robust subspace is learned from contaminated data and used for comparing the distance between higher moments  from robust estimate and moments of new observations.

\section{Experiments}
\label{sec:4_experiments}

This section presents the performed experiments and the acquired results. First, in Section \ref{sec:4_AnalyzedScenario}, the experimental scenario adopted in the experiments is summarized. Section \ref{sec:4_results} shows the results of the proposed approach, compares to selected anomaly detection algorithms, and presents a discussion about the results.

\subsection{Experimental Scenario}
\label{sec:4_AnalyzedScenario}

Anomaly detection algorithms usually rely on supervised or unsupervised methods, where the former requires labeled normal and anomalous data for training anomaly detection models, while the latter does not require labeled data or training. However, the availability of labeled data is a challenging concern in real world problems of anomalous detection, where anomalies are rare or even unkown events.

Semi-supervised algorithms are an alternative for the anomaly detection problem, considering that this method only relies on normal data for training and that non-malicious data can be obtained from historical information and from rule-based approaches. The proposed approach is semi-supervised and relies on normal data $\boldsymbol{X}^n$ and on contamination rate $c$ for training and anomaly detection, respectively.

Only for evaluation purpose, we use part of the contaminated data, denoted as $\textbf{X}^v$, for cross-validation, in order to estimate the contamination rate $c$ that best detect anomalies from $\textbf{X}^v$. However, we assume that the $c$ is well known or can be estimated for real world problems of anomaly detection.

Clustering methods for anomaly detection rely on three key assumptions: legitimate data instances often fall into a cluster whereas attacks do not, legitimate data instances are usually located near the closest cluster centroid while anomaly ones are often far away from it, legitimate data instances fall into vast and dense clusters and anomalies into small or spare ones \cite{ahmed2016survey, moustafa2019holistic}. Gaussian Mixture Model (GMM) is a statistical-based and parametric approach algorithm that estimates the distribution of the normal class from a training set and is typically based on a set of kernels \cite{moustafa2019holistic}.

This experimental evaluation compares our proposal to widely adopted algorithms for anomaly detection based on clustering and statistical approaches. As clustering-based method, we compare to K-means, which is the clustering algorithm most adopted for general purposes and for network anomaly detection \cite{gaddam2007kmeans,moustafa2019holistic}. It is observed that network traffic do not belong to a Gaussian distribution \cite{benson2010network,moustafa2019holistic}, and that algorithms based on non-Gaussian distributions, such as GMM, has been adopted for network anomaly detection \cite{moustafa2019holistic}. Therefore, we compare our proposal to GMM. We also evaluate the results of ROBPCA for anomaly detection on CTU-13, considering that ROBPCA also relies on robust estimates with adjusted outlyingness based on robust skewness \cite{hubert2009robustskewed}.

In contrast to Garcia \emph{et al.} \cite{garcia2014empirical}, we propose to evaluate each scenario of the CTU-13 data set individually, in a semi-supervised approach that does not rely on training data with labeled anomalies, in order to evaluate our proposed approach to all botnet malwares of the CTU-13.

Therefore, the full data set and scenarios can be denoted as $\boldsymbol{X} = \{\boldsymbol{X}_{10}, \boldsymbol{X}_{11}, \\\ldots , \boldsymbol{X}_{18-2}, \boldsymbol{X}_{19}\}$, in accordance to IDs presented in Table \ref{tab:4.01}. In our experiment setup each contaminated scenario $\boldsymbol{X}_i$ is split into $\boldsymbol{X}_i^n$ containing 50\% of the the normal data, and into the cross validation $\boldsymbol{X}_i^v$, composed by 25\% of the normal and 50\% of the anomalous data, and into test $\boldsymbol{X}_i^t$ containing 25\% of the normal and 50\% of the anomalous data.

In anomalous detection problems, where anomalies are rare events, if one classify all observations as normal and apply an accuracy evaluation, one would have high accuracy but poor true-positive detection. Due to the importance given by the F-measure to true-positive detection, it is the preferable measure for imbalanced data sets \cite{powers2011evaluation,moustafa2019holistic} and it is the metric used for validation of this experiment.

The F-measure is the harmonic mean of precision and recall and is defined by:
\begin{equation}\label{eq:4.15}
	F-measure = 2 * \frac{pr * rc}{pr + rc},
\end{equation}
where $pr$ is the precision, defined by 
\begin{equation}\label{eq:4.16}
	pr = \frac{True Positive}{True Positive + False Positive}
\end{equation}
and $rc$ is the recall, defined by 
\begin{equation}\label{eq:4.17}
	rc = \frac{True Positive}{True Positive + False Negative}.
\end{equation}

It is from $\boldsymbol{X}_i^n$ that we perform the training steps, which are the steps between 1 and 10 described in Algorithm \ref{alg:4.01}, resulting into $\boldsymbol{\bar{x}}$, $\boldsymbol{\bar{\Sigma}}$, $\boldsymbol{\bar{s}}$ and $\boldsymbol{\bar{k}}$. The next step of the experiment is the cross validation, which is an empirical testing to identify the contamination $c$ that obtains the best F-measure on anomaly detection, aiming the maximum precision and recall. During the cross validation for $\boldsymbol{X}_i^v$ we compute the steps 11, 12 and 13, and iterate over the steps between 14 and 16 for selected values of $c$, and validate the results of $\boldsymbol{\bar{t}}_{\bar{\boldsymbol{x}}}^v$, $\boldsymbol{\bar{t}}_{\bar{\boldsymbol{s}}}^v$, $\boldsymbol{\bar{t}}_{\bar{\boldsymbol{k}}}^v$ to the ground truth results for $\boldsymbol{X}_i^v$, in order to identify the $c$ that best classify the anomalies of $\boldsymbol{X}_i^v$.

Finally, we test the proposed approach to anomaly detection on $\boldsymbol{X}_i^t$ by means of the steps between 11 and 16 of Algorithm \ref{alg:4.01}, and compare the results of $\boldsymbol{\bar{t}}_{\bar{\boldsymbol{x}}}$, $\boldsymbol{\bar{t}}_{\bar{\boldsymbol{s}}}$, $\boldsymbol{\bar{t}}_{\bar{\boldsymbol{k}}}$ to the ground truth results for $\boldsymbol{X}_i^t$.

\subsection{Results}
\label{sec:4_results}

The Table \ref{tab:4.02} presents the results of the proposed approach and selected methods applied to botnet detection for each scenario of the CTU-13 data set. The Table \ref{tab:4.02} shows the F-measure of GMM, K-means, ROBPCA, md-RPCA, kd-RPCA and sd-RPCA by scenario and by an average of all scenarios.

\begin{table}[h!]
  \centering
  \scriptsize
  \caption{Results grouped by algorithm and scenario}
  \label{tab:4.02}
  \begin{tabular}{ c|c c|c c|c c|c c|c c|c c }
	\toprule
	\multirow{2}{*}{\textbf{Scenario}}   &\multicolumn{2}{c}{\textbf{GMM}} &\multicolumn{2}{c}{\textbf{K-Means}} &\multicolumn{2}{c}{\textbf{ROBPCA}} &\multicolumn{2}{c}{\textbf{md-RPCA}} &\multicolumn{2}{c}{\textbf{kd-RPCA}} &\multicolumn{2}{c}{\textbf{sd-RPCA}}\\ 
			\hhline{~------------}
			&\textbf{0.15} &\textbf{0.25} &\textbf{0.15} &\textbf{0.25} &\textbf{0.15} &\textbf{0.25} &\textbf{0.15} &\textbf{0.25} &\textbf{0.15} &\textbf{0.25} &\textbf{0.15} &\textbf{0.25}\\
	\midrule
		10 & 0.34 & 0.45 & 0.20 & 0.25 & 0.29 & 0.34 & 0.97 & 0.96 & 0.99 & 0.99 & 0.99 & 0.99 \\ \hline
		11 & 0.34 & 0.47 & 0.12 & 0.14 & 0.27 & 0.37 & 0.98 & 0.96 & 0.99 & 0.99 & 0.99 & 0.99 \\ \hline
		12 & 0.00 & 0.00 & 0.09 & 0.11 & 0.08 & 0.11 & 0.86 & 0.80 & 0.00 & 0.00 & 0.94 & 0.90 \\ \hline
		15 & 0.14 & 0.21 & 0.08 & 0.11 & 0.05 & 0.07 & 0.85 & 0.79 & 1,00 & 0.96 & 0.99 & 0.96 \\ \hline
		15-2 & 0.32 & 0.14 & 0.18 & 0.25 & 0.21 & 0.30 & 0.85 & 0.82 & 0.98 & 0.99 & 0.98 & 0.99 \\ \hline
		15-3 & 0.00 & 0.01 & 0.22 & 0.28 & 0.28 & 0.38 & 0.95 & 0.96 & 0.00 & 0.00 & 0.99 & 0.98 \\ \hline
		16 & 0.50 & 0.48 & 0.06 & 0.16 & 0.06 & 0.10 & 0.85 & 0.83 & 0.00 & 0.00 & 0.96 & 0.00 \\ \hline
		16-2 & 0.02 & 0.08 & 0.03 & 0.02 & 0.00 & 0.00 & 0.22 & 0.26 & 0.80 & 1,00 & 0.80 & 1,00 \\ \hline
		16-3 & 0.00 & 0.01 & 0.02 & 0.04 & 0.01 & 0.03 & 0.60 & 0.66 & 0.00 & 0.00 & 0.84 & 0.98 \\ \hline
		17 & 0.16 & 0.46 & 0.35 & 0.38 & 0.72 & 0.75 & 0.81 & 0.78 & 0.02 & 0.78 & 0.81 & 0.78 \\ \hline
		18 & 0.21 & 0.17 & 0.13 & 0.19 & 0.13 & 0.15 & 0.91 & 0.88 & 0.97 & 0.98 & 0.97 & 0.98 \\ \hline
		18-2 & 0.50 & 0.54 & 0.61 & 0.70 & 0.27 & 0.24 & 0.80 & 0.85 & 0.99 & 0.98 & 0.99 & 0.98 \\ \hline
		19 & 0.16 & 0.20 & 0.19 & 0.17 & 0.10 & 0.13 & 0.91 & 0.91 & 0.96 & 0.99 & 0.96 & 0.99 \\ \hline
		\rowcolor{Gray} \textbf{avg} &\textbf{0.21} &\textbf{0.25} &\textbf{0.18} & \textbf{0.21} & \textbf{0.19} &\textbf{0.23} &\textbf{0.81} &\textbf{0.80} &\textbf{0.59} &\textbf{0.67} &\textbf{0.94} &\textbf{0.89} \\
    \bottomrule
  \end{tabular}
\end{table}

It is possible to observe that the results of md-RPCA, kd-RPCA and sd-RPCA overcome the remain methods in a comparison by scenario and by average F-measure.

% 10: moment are better
% 11: moment are better
% 12: kd-RPCA is poor for scenario 12, as well as GMM, K-Means and ROBPCA
% 15: moment are better
% 15-2: moment are better
% 15-3: kd-RPCA is poor for , whilhe ROBPCA obtain 0.38 and other moment based has good results
% 16: kd-RPCA is poor for , as well as sd-RPCA with 0.25s, while GMM obtain 0.50 and other moment based has good results. Explain based on data
% 16-2: md-RPCA is poor for , as well as GMM, K-means and ROBPCA, but kd-RPCA and sd-RPCA goes well
% 16-3: kd_RPCA is poor, as well as GMM, K-means and ROBPCA
% 17: robpca has high result, higher than kd-RPCA, GMM, K-means, but md and sd are better and equal
% 18: moment based is better
% 18-2: high GMM e Kmeans, but moment are even better
% 19: moment are better
% there are no lots of researches using CTU-13, although it is been growing poor results for GMM, K-Means and MGM are ok and in accordance to \cite{wang2017botnet}, which also evaluate \cite{gu2007bothunter}

% a questão é diferenciar aquilo que é malicioso ou não, dentre os outliers ou novelties (esses termos são sinônimos para alguns autores)
% como eu posso diferenciar isto? Eu suponho que a partir de um modelo de tráfego normal eu possa detectar grandes deviations e classificar estas como ataques
% a hipotese é esta
% em outra área (compressive sensing), se utiliza dicionários, que são representações compactas de todas as caracteristicas de um dado
% no nosso caso, o treinamento nos gera uma especie de dicionário, em que os dados normais são representados de forma robusta
% onde uma forma robusta seria uma forma menos sensível a outliers. Isso pode ser visto como a diferenca entre calcular a média e a mediana de um conjunto. A presença de um outlier no conjunto irá corromper a média de uma forma maior que a mediana. Desta forma, se considera a mediana como sendo robusta

\section{Results}
\label{sec:4_results}

\section{Conclusion}
\label{sec:4_conclusion}

This paper models the network traffic as a signal processing formulation for applying to the framework for detection and identification of network attacks, which is based on eigenvalue analysis, model order selection (MOS) and eigen similarity analysis.

The proposed framework is evaluated and the experimental results show that synflood, fraggle and port scan attacks can be detected accurately and with great detail in an automatic and blind fashion, applying signal processing concepts for traffic modeling and through approaches based on MOS and eigen similarity analysis. The main contributions of this work were: the extension of an approach based on MOS combined with eigen analysis to blindly detect time frames under network attack; the proposal and evaluation of an eigen similarity based framework to identify details of network attacks, presenting F-measure of timely detection and identification of TCP/UDP ports under attack, as well as presenting acceptable complexity and performance regarding the processing time.

This paper evaluated the effectiveness of MOS schemes for fraggle attack detection, extending our previous work and showing that the analysis of the largest eigenvalues by time frames can be applied to detect the number of port scanning, and flood attacks, but still requiring more information for detailed attack detection. Therefore, we proposed a novel approach for detailed network attack detection, based on eigen similarity analysis.

The incremental individualized approach of eigen similarity analysis, is able to detect low similarity for all evaluated scenarios and types of network attack, while the other approaches present false positives or low sensibility to eigen similarity analysis for network attack detection. Therefore, the incremental individualized approach is able to gradually and incrementally adapt to network traffic changing, preserving the sensibility to identify outliers or anomalies by time or network port, and reducing the occurrence of false positives.

According to the significant similarity difference between legitimate and malicious traffic, it is possible to adopt safe thresholds for flood and port scan detection through eigen similarity analysis.

Future research is directed to improvements for obtaining better false positive rates, as well as for make the proposed framework able to identify sparse probe attacks or subtle behaviors, such as exfiltration or covert communication, considering the evaluation of a flow-based analysis and novel data sets. Distributed or parallel processing can also be evaluated to analyze the scalability and processing capacity for monitoring02       rch can evaluate the application of the proposed approach to different attack types and domains, considering cases that are aware to behavioral analysis.

Test distribution fit again e present results 
Conclude this presentation.
Evaluate eigensim for CTU-13 and others algorithms for Danilo's data set.
Compare to vanilla RPCA (Eduardo is working on it).
Write a paper for JNCA (April).
Write the thesis (May and June).

say that more studies about skewness and subspace learning can be done in order to explore it inside algorithms and non gaussian data

note that the semi supervised approach requires subspace learning only for training and not for testing, while unsupervised requires subspace learning for each evaluations