### 2019-06-16 docs/thesis/thesis.tex

%% These files were modified in Lyx by:
%           Lus F. C. Fgueiredo, Fev., 2017.
%% Git repo for Lyx version are available at:
%%          https://github.com/lara-unb/DOCUMENTS/tree/master/templates
%% Latex version for sharelatex and overleaf v2 ported by:
%%          Vinícius Galvão Guimarães
%% 
%% Contributions of:
%%          Miguel Eduardo Gutierrez Paredes
%% LyX 2.2.3 created this file.  For more info, see http://www.lyx.org/.

%\documentclass[11pt,oneside,portuguese]{book}
\documentclass[12pt,oneside,english]{book}
\input{setup/packages}
\input{setup/other}
\input{setup/theorem_lemma}
\input{setup/languages}
\input{setup/listing_def}
\input{setup/glossaries_setup}

\makeatother
\begin{document}
\pagenumbering{roman}

\input{setup/title_author_etc}
\input{parts/dedicated_to}

\runfrontend{}

\import{}{macros}
\import{parts/}{abstract}

\sumario

\listadefiguras

% \listadetabelas

%-----------------------------------------------
% Use os simbolos, acronimos e notacoes com as funcoes a seguir
% To use this acronyms list, check the file glossaries-list/symbos.tex.
	\listadesimbolos

% To use this acronyms list, check the file glossaries-list/acronyms.tex.
	\listadeacronimos

% To use this acronyms list, check the file glossaries-list/notation.tex.
	\listadenotacoes
%-----------------------------------------------

\markboth{}{}

%Main Body
\mainmatter 
\setcounter{page}{1} \pagestyle{plain}

\import{parts/}{1_introduction}
\import{parts/}{2_eigen_similarity}
\import{parts/}{3_m_rpca}
\import{parts/}{4_mobile}
\import{parts/}{5_conclusion}

%BIBLIOGRAPHY
\addcontentsline{toc}{chapter}{REFERENCES}
% Do not use \bibliographystyle{IEEEtran} with abntex2cite package!!!!!
% Also, try not to use abntex2cite at all! just comment its package in setup/packages.tex
%\bibliographystyle{chicago}
%\bibliographystyle{acm}
\bibliographystyle{IEEEtran}

\bibliography{bibliography}

%APPENDIX
\part*{APPENDIX}
\appendix
\import{parts/}{a_mos}
\import{parts/}{b_csf_pca_svm}

\end{document}